\chapter{Discussion}
\label{ch:discussion}

The discussion is divided into four sections.  First, interpretation and
significance of the results -- including some comment on experimental methods
which were not used in the final analysis.  Second, reflections on the methods
which \emph{were} used, and on the limitations of this research from both
scope and methodology.  Third, an attempt to assess potential applications or
benefits of this research.  Finally, several suggested directions for future
research in this area.



\section{Discussion of Results}
\label{sec:disc-results}

\subsection{Qualitative and Interview Results}

Interviews and informal discussions are an essential part of this study --
that much was obvious in advance, and easily justified the cost of fieldwork.
In retrospect they were indeed essential, for unforseen reasons!
%
Two goals necessitated fieldwork and in-person discussion of Yolngu seasons.
The first was to collect detailed descriptions of each season, the definitions
and timing, extreme events, climate change, and so on; the qualitative results
show this was a success.  However, this is in a sense preliminary work --
enough to get exploratory results, that can be taken back to Indigenous
research participants and partners to start a deeper discussion.  In future
studies secondary sources such as \textit{Man of All Seasons} \citep{davis1989}
may support this preliminary analysis, and allow researchers to demonstrate
their intentions, at some risk of wasted effort.

The second goal of fieldwork  was to learn about the structure of Yolngu
seasons and try to catch any important information that would otherwise
fail to cross cultural barriers.  The volume of \cref{sec:complex-seasons},
on \textit{\nameref{sec:complex-seasons}}, speaks to the importance of these
results.
%
This validated the methodology -- without direct conversations, or if I had
assumed that published material set out all relevant information, crucial
information be missing and the results would only be of any use to the extent
that their accurate portrayal of Yolngu knowledge was irrelevant.
%
In short: there is little to no value in quantitative or numerical analyses
founded on faulty premises, no matter how rigorous or technically impressive.
An investment in cross-cultural understanding is well worth making, especially
on complex topics that appear to be simple!



\subsection{Weather Observations}

The observational weather record, shown in
\cref{fig:galiwinku-observations,fig:milingimbi-observations}, displays
several interesting patterns.  Each variable has a distinct seasonality,
the timing and in some cases intensity of which visibly varies between
years.  The dates of first and last rainfall vary by months; in some
years rain falls steadily and in others all at once.
%
Minimum and maximum temperatures follow slightly different patterns at
Galiwinku to those at Milingimbi -- for example, while both record their
highest temperatures around November, the following months are consistently
cooler at Galiwinku.

Wind in NE Arnhem Land appears to be dominated on a seasonal basis by the
monsoonal winds, and at these locations is hevily influenced by the sea
breeze.  Background conditions can be seen in 9am wind, which is minimally
affected by the sea breeze (see \cref{fig:galiwinku-seabreeze-direction}
for three-hourly data; speed and other stations in electronic appendices)
and displays a westerly or south-easterly regime depending on the monsoon.
Morning wind speed is steady throughout the year; at Galiwinku heavy rain
is associated with stronger winds.
%
At 3pm, things are more complicated.  The sea breeze blows from the north
for most of the year; but shows a south-easterly influence from May to
August.  Given that this is stronger at Galiwinku, which is around 80km
east of Milingimbi, I attribute this to a combination of the monsoon wind
and the sea breeze off the Gulf of Carpentaria, blowing across the intervening
land.  Speeds are similar throughout the day at Milingimbi, while Galiwinku
peaks in the afternoon.



\subsection{Detection of Seasons}
\label{subsec:disc-season-detection}

\todo{
Note that trend- and event-based approaches were tested and didn't work.
Discuss how well season detection worked, nondistinctiveness of wet
season.  Propose algorithm with memory to fix this.  (and implement?)
Really hammer home the non-european-ness of these seasons,
including by listing the shortcuts that would not be well supported.
Discuss various ways to characterise onset and typical timing, issues with each,
foreshadow the final plot for conclusions.
}


\subsection{Quantitative Analysis}

\todo{discuss this stuff}



\section{Reflection - methods and limitations}
\label{sec:disc-reflection}

\todo{What were the methods?  How well did they work?
Are they applicable to other calendars?  If so, to what extent?
}





\todo{Write in general terms about what -- of the in-scope things --
were limitations or didn't work.  Where did scoping hurt?
What could or should have been done better?  How much was foreseeable?
Detailed critique belongs in specific sections above, though.
}



\section{Applications or Benefits of the Research}
\label{sec:applications-benefits}

Thinking about novelty here.  Also relevance to people's lives,
industry, etc.  Argue that I'm contributing to the ongoing process
of non-Indigenous Australians adapting to this country.

\todo{Still missing an elucidation of why the research matters to Yolngu -
and everyone else!}
Looking at abstract benefits (contribution to knowledge, valuing Indigenous knowledge),
but also cover concrete benefits - what activities in northern australia
in particular (eg land and NR management) will benefit from these insights?
Key aspect: how transferable are these insights?

~\\

Seasons provide powerful heuristics for the expected range and variability in
weather for a given time of year.

-	Implications of seasonality, relationship to NRM, applications, meaning of
change, what does this way of thinking translate into action


Quantitative knowledge of local indigenous seasonality can improve practises in
natural resource management, fire control, agriculture, and tourism -
particularly outside the temperate zone.  Seasons can provide powerful
heuristics for the expected range and variability in weather, and may yield
insights into patterns of change.  For practical use, a local and long-tested
seasonal calendar has substantial advantages over one imported from Europe.


See Barber 2005 ``where the clouds stand'' (about Tiwi Islands)

This research is expected to make a novel contribution to knowledge, raise
awareness of Yolngu culture, and improve understanding of the climate
seasonality of NE Arnhem Land.  The methods demonstrate two-ways research in
combining indigenous knowledge with numerical data to characterise seasons.

For non-technical communities, both Indigenous and non, discussing climate
change in terms of seasonality is more meaningful than degrees of warming.
Framing the impacts of climate change in terms relatable to lived experience
has the potential to boost engagement with and understanding of the major
results of climate science - after all, everyone has some experience of the
weather!

For the Yolngu community, increased engagement with the science around climate
impacts has benefits for both adaptation choices and advocacy for mitigation,
which are particularly important for coastal communities living off the land.
Yolngu have expressed strong interest in other records of and perspectives on
past environmental changes.

For non-indigenous Australians, this research contributes to the continuing
process of learning to live on this continent, discarding `maladaptive traditions'
- eg our poor infrastructue planning for drought and fire (REF the future eaters).



\section{Directions for Further Study}
\label{sec:further-study}

Having demonstrated the viability of quantitative analysis of Indigenous
seasons, and discusses the value of such research, there are three distinct
directions for further study:  increasing one or more of the breadth, depth,
or extent of the research.  Alternative topics may also address the aim
of this research -- \textit{``developing an approach to integrating traditional
knowledge with a quantitative analytical approach''} -- but are not suggested
here.


\subsection{Breadth of Study}

Increasing the breadth of study would result in similar but less exploratory
research; essentially trading novelty for rigor.  Generally this means
collecting and analysing more data.

\begin{description}
\item[Multiple Indigenous calendars]
    Replication of this study with other Indigenous calendars would be
    valuable in each case, and the ability to compare calendars and investigate
    possible correlations between local climate and how seasons are defined
    could yield fascinating insights.

\item[Single calendar, multiple sites]
    Similarly, comparing the calendar of a single language group across
    multiple sites and subtly different climates could be very interesting.
    An `objective' weather record could really help researchers to draw out
    the differences between calendars, and perhaps see some of the subtleties.
    While this study had participants and weather observations from multiple
    sites, it did not analyse them as qualitatively distinct records or
    attempt any comparisons.

\item[Iterative fieldwork]
    Future work should if at all possible invest the time and resources
    neccessary for multiple visits to Indigenous homelands.  Second opinions,
    feedback on quantitative results, and the strong relationships that
    long-term engagement can grow could all prove surprisingly valuable.

\end{description}


\subsection{Depth of Study}
Increasing the depth of study would \todo{elaborate}


\begin{description}
\item[Kinds of seasons]
    \todo{Comparing the various kinds of seasons within a calendar would be really interesting}

\item[Ecological knowledge]
    Investigating Indigenous ecological seasons could provide deep insights
    into local ecology and traditional natural resource management.
    The principle challenges are resourcing, to collect the quantitative
    data required to `translate' back to western science, and lengthy
    investment in time and relationships required.

\item[Per-year patterns]
    \todo{(rephrase) Look at patterns of interuptions within single years.  Are there distinct
    kinds of seasonal occurance that a year may exhibit?}

\item[]

\end{description}



\subsection{Extent of Study}
Increasing the extend of study would mean expanding the topic; perhaps by
looking at links to the wider context -- climate indicies or long-term
change, non-Indigenous calendars, and so on.

\begin{description}
\item[Non-Indigenous calendars]
    \todo{something}

\item[Other weather data]
    The methods used in this study can be applied to any dataset which
    includes a daily record of the relevant variables.  This enables
    investigation of changes over time -- past or future -- and scenarios
    including a wider range of conditions than exist in the historical record.

\item[Climatology]
    \todo{A proper look at the climatology of Indigenous seasons would
    be cool, especially looking at the relationship with non-surface weather}

\item[Large-scale indices]
    \todo{Detailed investigation of monsoon correlations, ENSO, IOD, etc}

\item[Climate Change]
    \todo{Will the seasons change?  And if so, how?}
    I plan to conduct and publish such a study using forecast and backcast
    model outputs from the ACCESS and CMIP5 ensembles immediately
    after graduating.

\end{description}

